The ANthropological Notation Ontology (ANNO): A core ontology for annotating human bones and deriving phenotypes

### 2023-12-19 paper.tex
add statistical information

% add. options: [seceqn,secthm,crcready]
\documentclass[sw]{iosart2x}
%\usepackage[T1]{fontenc}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{csquotes}
\usepackage{aurl}
\usepackage{booktabs}
\usepackage{tabulary}
%\usepackage{dcolumn}
%\usepackage{placeins} %provides \FloatBarrier

@@

The resulting OWL 2 ontology was generated using SMOG [...] and validated by the HermiT reasoner [...] and SHACL shapes (chapter $\ldots$).
In both cases, the consistency of the ontology was proven.
As a proof of concept, we integrated the ontology into the AW3D software [...] and were able to successfully realise all three intended use cases.
Thereby, most of the core categories (such as \enquote{Bone}, \enquote{Bone part}, \enquote{Bone compound}, \enquote{Tooth}, \enquote{Tooth part}, \enquote{Phenotype}, \enquote{Relative anatomical location}, \enquote{Anatomical line} and \enquote{Anatomical point}) and relation types (such as hasPart, boundaryOf, locationOf and derivedFrom) of ANNOdc were utilised and their subcategories and specific relations were dynamically queried (according to the three-ontology method).

@@

\begin{table}[b]
\centering
\caption{Statistical information.}
\label{tab:stats}
\begin{tabulary}{\textwidth}{lllll}
\toprule
Ontology	&Classes	&Object properties	&Annotation properties\\
\midrule
ANNOdc\\
ANNOds\\
\bottomrule
\end{tabulary}
\end{table}

>>>>

% add. options: [seceqn,secthm,crcready]
\documentclass[sw]{iosart2x}
%\usepackage[T1]{fontenc}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{csquotes}
\usepackage{aurl}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{siunitx}
%\usepackage{dcolumn}
%\usepackage{placeins} %provides \FloatBarrier

@@

The resulting OWL 2 ontology was generated using SMOG [...] and validated by the HermiT and Pellet reasoners [...] and SHACL shapes (see \cref{sec:domain}).
In both cases, the consistency of the ontology was proven.
As a proof of concept, we integrated the ontology into the AW3D software [...] and were able to successfully realise all three intended use cases.
Thereby, most of the core categories (such as \enquote{Bone}, \enquote{Bone part}, \enquote{Bone compound}, \enquote{Tooth}, \enquote{Tooth part}, \enquote{Phenotype}, \enquote{Relative anatomical location}, \enquote{Anatomical line} and \enquote{Anatomical point}) and relation types (such as hasPart, boundaryOf, locationOf and derivedFrom) of ANNOdc were utilised and their subcategories and specific relations were dynamically queried (according to the three-ontology method).

@@

\begin{table}[b]
\centering
\caption{Statistical information. Column values do not add up because some entities are defined in both subontologies.}
\label{tab:stats}
\begin{tabulary}{\textwidth}{lSSSSS}
\toprule
Ontology	&\textnormal{Classes}	&\textnormal{Object properties}	&\textnormal{Individuals}	&\textnormal{Annotation properties}	&\textnormal{Axioms}\\
\midrule
ANNOdc		&41			&18					&0				&19						&178\\
ANNOds		&1166		&11					&803			&43						&18231\\
\midrule
combined	&1195		&21					&803			&53						&18382\\
\bottomrule
\end{tabulary}
\end{table}
